---
title: "LA.5 Tutorial"
output:
  learnr::tutorial:
    theme: "united"
runtime: shiny_prerendered
description: >
  Learn the basics of `ggplot2` for graphing.
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(descr)
library(magrittr)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction
This tutorial covers the basics of plotting with the package, `ggplot2`. This package is contained within a collection of R packages designed for data science known as the [tidyverse](https://www.tidyverse.org/).

The core tidyverse includes packages that we use in everyday data analysis, including:

- `ggplot2`
- `dplyr`
- `tibble`

In this tutorial, we will learn how to:

1. Create a histogram
2. Create a scatter plot

Before you begin this tutorial, read [Chapter 5 of COMM 3710: Getting Started with R](https://sarakyeo.github.io/COMM-3710-bookdown/docs/ggplot2.html).


## Histograms
A histogram is a graphical representation of a frequency distribution of a variable. It shows the values of a variable and the number of units in your sample that are assigned that value. You have already seen many histograms; the `freq()` command in the `descr` that we have been using automatically generates a frequency table and a histogram showing the distribution of the variable.

Although `freq()` generates a histogram, we cannot easily modify it (e.g., adding axes labels, legends). To do this, we can use `ggplot2` to create a histogram, among many other types of graphs. We will use the `mpg` dataset again. To refresh your memory, open the help page for `mpg` using the code chunk below.

### Example
Let's see what `ggplot2` is doing when you create a histogram with `ggplot2`. Run the code chunks below one at a time and observe the output to get a better sense of how ggplot builds graphs in layers.

```{r plot1, exercise=TRUE}
ggplot(data = mpg)
```

```{r plot2, exercise=TRUE}
ggplot(data = mpg, aes(x = hwy))
```

```{r plot3, exercise=TRUE}
ggplot(data = mpg, aes(x = hwy)) +
        geom_histogram()
```

We can add axis labels to our graph by using the `labs()` layer.

```{r labels, exercise=TRUE}
ggplot(data = mpg, aes(x = hwy)) +
        geom_histogram() +
        labs(x = "Highway gas mileage (miles per gallon)")
```

### Your turn
Create a histogram to examine the values of `cty`. Label the x-axis with "City gas mileage (miles per gallon)."

```{r student-hist, exercise=TRUE}

```

```{r student-hist-hint}
ggplot(data = mpg, aes(x = cty)) +
        geom_histogram() +
        labs(x = "City gas mileage (miles per gallon)")
```


## Scatter plots
Now, let's examine the relationship between `cty` and `hwy`. What do you think this relationship might be? We can check our hypothesis about the relationship between these variables by creating a scatter plot.

To do so, we specify both the x and y variables in the `aes()` argument of the `ggplot()` function. In the histogram, we specified `x = hwy` or `x = cty`. In this case, we also need to include a `y = ...` in the argument.

In addition to specifying which variable to plot on the y-axis, we also need to tell `ggplot2` that we want a scatter plot. We do so by changing `geom_histogram()` into `geom_point()`.

### Your turn
Use the code chunk below to create a scatter plot with `hwy` on the x-axis and `cty` on the y-axis. Label both axes appropriately.

```{r scatter, exercise=TRUE, exercise.lines=10}

```

```{r scatter-hint}
ggplot(data = mpg, aes(x = hwy, y = cty)) +
        geom_point() +
        labs(x = "Highway gas mileage (miles per gallon)",
             y = "City gas miileage (miles per gallon)")
```

### Congratulations
You are now ready to complete LA.5!
