---
title: "LA.3 Tutorial"
output:
  learnr::tutorial:
    theme: "united"
runtime: shiny_prerendered
description: >
  Learn how to access help pages, run frequency distributions, and subset data
  in R.
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(descr)
library(magrittr)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction
In this tutorial, you will learn how to:

1. Open a help page
2. Run a frequency distribution
3. Subset data


## Data frames
### What is a data frame?
A data frame is a rectangular collection of values, usually organized so that variables appear in the columns and observations appear in rows. For this tutorial, we will use the `mpg` data frame that is included in one of the R packages that we will be using in this course, `ggplot2`.

The `mpg` data frame contains observations collected by the U.S. Environmental Protection Agency on 38 models of cars. To see the `mpg` data frame, type `mpg` in the code chunk below and then click “Run code.”

```{r df, exercise=TRUE}

```

### A note about `mpg`
Note that I have already loaded the `ggplot2` package for you. On your own computer, you will first have to install and load `ggplot2` in R.

To do so, follow the instructions in this [section of COMM 3710: Getting Started with R](https://sarakyeo.github.io/COMM-3710-bookdown/docs/start.html#packages).


## Help pages
To learn more about a command or function in R, you can open its help page. The help page explains the functions, its arguments, and includes use examples. Be sure to review the examples, which are often at the bottom of the help page.

Let's learn more about the `mpg` dataset by opening its help page. Type `?mpg` in the code chunk below and click "Run code."

```{r help, exercise=TRUE}

```

The page that opens in your web browser gives you a description of the `mpg` dataset as well as information about the number of observations (rows) and variables (columns).


## Frequency distributions
A frequency distribution is an overview of all distinct values in a variable and the number of times they occur. A frequency distribution can be shown as a list, table, or graph.

The definition below uses terms with which we are familiar from lecture:

> A frequency distribution is a description of the number of times the various attributes of a variable are observed in a sample.

Let's learn how to run frequency distributions in R using the `freq()` function in the `descr` package. I have already loaded the package for you, but you will have to do so when working on your assignments for this course.

### Example
The R code below runs a frequency distribution of the variable, `manufacturer`, in the `mpg` data frame. Running the code produces a **histogram** and a **frequency table**.

```{r example-freq, exercise=TRUE}
freq(mpg$manufacturer)
```

### Your turn
Write the R code to run a frequency distribution of the variable `hwy` in the `mpg` data frame.

```{r student-freq, exercise=TRUE}

```

```{r student-freq-hint}
freq(mpg$hwy)
```


## Subset data
Often, we may want to examine a subset of our data. To do so, we can use the `filter()` function. 

### Example
Let's say I want to examine the highway gas mileage (`hwy`) of only compact cars. To do so, I can subset the data by the variable, `class`, and then run a frequency distribution of `hwy`.

The code below subsets the data based on the variable, `class`, into a separate data frame called `subdf` (you can call this whatever you want). The subsequent line of code runs a frequency distribution on the variable `hwy` in the `subdf` data frame. 

```{r example-sub, exercise=TRUE}
subdf <- filter(mpg, class == "compact")
freq(subdf$hwy)
```

### Your turn
Write R code to subset the data by manufacturer (`manufacturer`). Examine the highway mileage (`hwy`) of cars made by Toyota by running a frequency distribution.

```{r student-sub, exercise=TRUE}

```

```{r student-sub-hint}
subdf <- filter(mpg, manufacturer == "toyota")
freq(subdf$hwy)
```

### Congratulations
You are now ready to complete LA.3!
