---
title: "LA.4 Tutorial"
output:
  learnr::tutorial:
    theme: "united"
runtime: shiny_prerendered
description: >
  Learn how to recode a variable using `ifelse()` and create a new variable
  from existing variables.
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(descr)
library(magrittr)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction
In this tutorial, you will learn how to:

1. Use `[ ]` for subsetting to recode variables
2. Use the `ifelse()` function to recode variables
3. Create a new variable using mathematical operations

We will use the `mpg` dataset from `ggplot2`, which has already been loaded for you in this tutorial. If you need to review the `mpg` dataset, open its help page (`?mpg`) in the code chunk below.

```{r help, exercise=TRUE}

```


## Recoding variables
**Recode** is a term used to describe the process of making changes to the values of a variable. Recoding a variable often involves changing values in an existing variable to make a new one.

Let's say we want to create a new variable called `citygm` that categorizes cars based on their city gas mileage, `cty`. We will have cars with 15 miles per gallon or lower be categorized as `low` and cars with over 15 miles per gallon be categorized as `high`.

### Subsetting to recode variables
The first method we can use to recode an existing variable into a new one is to use square brackets, `[ ]` for subsetting.

We can use `[ ]` to select any number of elements from a vector. In this case, we will use a **logical vector** to select elements from the original variable, `cty`. The symbols `<=` and `>` are [logical operators](https://sarakyeo.github.io/COMM-3710-bookdown/docs/start.html#logical-operators).

Before we begin the recoding process, we should examine the frequency distribution of the variable we want to recode, `cty`.

*Use the code chunk below to run a frequency distribution on `cty`.*

```{r freq-1, exercise=TRUE}

```

```{r freq-1-hint}
freq(mpg$gm)
```

Now that we have examined the values of `cty`, we can recode it. The code below selects elements in `cty` that are equal to or lower than 15 and assigns the value `low` to them in the new variable `citygm`. The next line of code selects elements in `cty` that are greater than 15 and assigns `high` to them in `citygm`.

*Run the code below to create `citygm`. Then, write code to run a frequency distribution of the newly created variable.*

```{r ex-square, exercise=TRUE, exercise.lines=10}
mpg$citygm[mpg$cty <= 15] <- "low"
mpg$citygm[mpg$cty > 15] <- "high"

## Write code below to run a frequency distribution

```

### Your turn
*Write R code to create a new variable, `hwygm`, that categorizes the cars based on highway gas mileage, `hwy`. Have the value of `hwygm` be `low` when `hwy` is less than or equal to 20 and `high` when `hwy` is greater than 20.*

Remember to run a frequency distribution of both the existing and new variables to check that you created correctly.

```{r student-square, exercise=TRUE, exercise.lines=10}

```

```{r student-square-hint}
## Check var to be recoded
freq(mpg$hwy)

## Recode hwy var into hwygm
mpg$hwygm[mpg$hwy <= 20] <- "low"
mpg$hwygm[mpg$hwy > 20] <- "high"

## Check new/recoded var
freq(mpg$hwygm)
```


## Using `ifelse()` to recode variables
Another means of recoding a variable is to use the function, `ifelse()`. First, use the code chunk below to open the help page for the function and read the documentation.

```{r student-help, exercise=TRUE}

```

```{r student-help-hint}
?ifelse
```

### Example
To begin, run a frequency distribution of the existing variable, `cty`.
The code below creates a new variable called `citygm` based on the existing variable, `cty`. It recodes `cty` into a new variable that assigns "low" to cars that have a city gas mileage of 15 miles per gallon or lower and "high" to cars that have higher than 15 city miles per gallon.

```{r ex-ifelse, exercise=TRUE, exercise.lines=10}
## Check var to be recoded
freq(mpg$cty)

## Recode cty into citygm using ifelse()
mpg$citygm <- ifelse(mpg$cty <= 15, "low", "high")

## Check new/recoded var
freq(mpg$citygm)
```

### Your turn
*Use the `ifelse()` function to create, `hwygm`. Have the value of `hwygm` be `low` when `hwy` is less than or equal to 20 and `high` when `hwy` is greater than 20. Remember to use `freq()` to check the existing and new variables.*

```{r student-ifelse, exercise=TRUE, exercise.lines=10}

```

```{r student-ifelse-hint}
mpg$hwygm <- ifelse(mpg$hwy <= 20, "low", "high")
```


## Mathematical operations to create a new variable
We can also create a new variable by summing/adding, subtracting, or calculating the average of two or more variables.

### Example
Let's create a new variable, `sumgm`, by adding the two variables, `cty` and `hwy`. Write R code to check the frequency distribution of the new variable, `sumgm`.

```{r ex-oper, exercise=TRUE, exercise.lines=10}
mpg$sumgm <- mpg$cty + mpg$hwy

## Write code below to run frequency distribution

```

### Your turn
It does not make much sense to add `cty` and `hwy` together. It would be more informative to know each car's average gas mileage. To do this, we can take an average of the two variables by adding them together and dividing by 2.

*In the code chunk below, create the `avgm` variable by averaging `cty` and `hwy`. Then, run a frequency distribution of `avgm` to check that you have created it correctly.*

```{e student-oper, exercise=TRUE, exercise.lines=10}

```

```{r student-oper-hint}
mpg$avgm <- (mpg$hwy + mpg$cty)/2
freq(mpg$avgm)
```

### Congratulations
You are now ready to complete LA.4!
